Reconstructing the Lensing Mass in the Universe \\ from Photometric Catalog Data

### 2012-10-31 doc/paper-lcr.tex
Body of text done

\author[Collett \etal]{%
  Thomas~E.~Collett$^{1}$\thanks{\collettemail},
  Philip~J.~Marshall$^{2}$,
  Matthew~W.~Auger$^{1}$,
  Stefan~Hilbert$^{3}$,
\newauthor{%
  Sherry~H.~Suyu$^{4}$,
  Zachary~Greene$^{4}$,
  Tommaso~Treu$^{4}$\thanks{\packard},
  Christopher~D.~Fassnacht$^{5}$,}
\newauthor{%
  L\`eon~V.~E.~Koopmans$^{6}$,
  Roger~D.~Blandford$^{3}$} 
  \medskip\\
  $^1$\ioa\\
  $^2$\oxford\\
  $^3$\kipac\\
  $^4$\ucsb\\
  $^5$\davis\\
  $^6$\kapteyn
}

@@

Real attempts to reconstruct the convergence along a line of sight will not have direct access to the masses of dark matter halos, 
and will likely not have access to a spectroscopic redshift either. Instead these properties must be inferred from astrophysical
observables. In principle the only observables may be astrometric positions, photometric colours and possibly spectra. In this section we
quantify the errors induced by inferring the halo mass and redshift from the observables. Much work has already focused on using
photometric colours to infer stellar mass \citep[\eg][]{AugerEtal2009} and redshifts \citep[\eg][]{BPZ}, in line with these we shall use
stellar mass and photometric redshifts (with appropriate errors), as the observables that must be converted into an estimate
of $\kappax$. We will investigate two main sources of error: inferring halo mass given observed stellar mass and placing halos at the wrong redshift due to photometric redshift error.

We generate a stellar mass for each halo according to the stellar mass--halo mass relation of 
\citet{BehrooziEtal2010}. From this stellar mass we simulate 
an observed stellar mass by drawing a sample from $\pr(\log(M_{* \mathrm {obs}})|\log(M_{* \mathrm {true}}))$ which is given by a Gaussian of width $\sigma_{M_*}$ centred on $\log(M*_{\mathrm {true}})$. Where a spectroscopic redshift exists stellar masses can be estimated with typical uncertainties of 0.15 dex \citep{AugerEtal2009}, however with photometric redshifts stellar mass uncertainties are typically three times as large; we use $\sigma_{M_*}=0.15$ for halos with a spectroscopic redshift and $\sigma_{M_*}=0.45$ otherwise. For photometric redshift errors we draw a redshift from $\pr(z_{\rm true}|z_{\rm obs})$ which we take as a Gaussian of width $0.1(1+z_{\rm spec})$ centred on $z_{\rm spec}$, where $z_{\rm spec}$ is the halo's true redshift in the \MS catalogue.

Given an uncertain stellar mass and redshift it is possible to infer a
halo mass using the stellar mass--halo mass relation of
\citet{BehrooziEtal2010}. This \proceedure requires an inversion of the
relation given in \citet{BehrooziEtal2010} and correctly inverting the
relation's uncertainties requires care: the \proceedure we use to do this
is given in Appendix \ref{appendix:MSMH}. By drawing sample halo masses
and redshifts, we can infer a sample $\kappah$ using the \proceedure of
Section \ref{sec:model}; repeatedly drawing samples allows us to
estimate P$(\kappah)$ for each reconstructed line of sight.

@@

\begin{figure}
\includegraphics[width=\columnwidth]{figs/Width.eps}
\caption[magcut]{The widths of the \infered P$(\kappah|\mathcal{D})$ for $10^5$ lines of sight given different amounts of data. Blue assumes a spectroscopic redshift for every halo with $i<26$, purple assumes a spectroscopic redshift for every halo with $i<23$ and every within 1 arcminute and $i<24$, cyan assumes a spectroscopic redshift for every halo with $i<22$ and green assumes that no objects in the field have a spectroscopic redshift}
\label{fig:reconwidths}
\end{figure}

\begin{figure}
\includegraphics[width=\columnwidth]{figs/Widthvshilberrt.eps}
\caption[magcut]{Width of the \infered P$(\kappah|\mathcal{D})$ verses
the true $\kappa$ along a line of sight. Blue assumes a spectroscopic
redshift for every halo, whilst green assumes no spectroscopic
redshifts. The width of the \infered P$(\kappah)$ increases with $\kappa$. If correctly reconstructed, the lowest density lines of sight can provide the most precise measurement of the time delay distance.}
\label{fig:widthsvsH}
\end{figure}

The best possible reconstruction of $\kappah$ corresponds to having a sepctroscopic redshift for every single object in our field. Observationally such a reconstruction would be prohibitively expensive in terms of telescope time, but we investigate this scenario as an ideal case. Reconstructing the lines of sight with perfect knowledge of the redshift but an uncertain stellar mass, we find that the width of $\Pr\left(\sumkappah \right)$ grows with the expectation value of $\sumkappah$; this is not surprising since the low $\sumkappah$ lines of sight are relatively empty and so there are few opportunities for uncertainties in the halo masses to propogate into $\sumkappah$ uncertainties. After applying the void correction the median reconstruction width is 0.0123; with 22 percent of lines of sight having a reconstruction width less than 0.01.

@@

Inferring the stellar mass of a galaxy from its magnitude and colours requires an estimate of how far away the galaxy is; without a spectroscopic redshift the the \infered stellar mass is less precise. However, obtaining photometry has much lower observational cost; over the next few years several large area photometric surveys will reach 24th magnitude \citep{EUCLID,LSST}, these will provide sufficient data to reconstruct lines of sight {\it without} additional observations. In principle the photometric redshift is correlated with the \infered stellar mass, however we do not model this effect since the convergence from the outskirts of an individual halo is only weakly dependant on redshift at fixed mass: the redshift error is small effect on the recovered $\pr(\kappa)$ when compared to the effect of uncertain stellar masses.  With only photometric redshifts the uncertainty on $\sum\kapparec^{\rm halos}$ is much larger than the spectrocopic case and this propogates into a broader $\pr(\kappa)$, although the photometric reconstruction still typically produces a 1.4 fold improvement compared to the precision of the global P($\kappa$).

@@

Whilst a fully photometric reconstruction provides useful constraints on $\pr(\kappa)$, targeted spectroscopy can provide addtional constraints upon the masses and redshifts of halos whose $\kappah$ contribution have the largest absolute uncertainty. Obtaining spectra of bright ($i<22$) objects is relatively fast with 8-m class telescopes; we find that if spectroscopic redhsifts were known for all $i<22$ galaxies in our field the reconstruction improves slightly upon the purely photometric reconstruction, although the improvement depends strongly upon the nature of the particular line of sight. Obtaining spectra for fainter objects would be considerably more expensive, but if spectroscopic redshifts could be obtained for all $i<23$ galaxies and all $i<24$ galaxies within 1 arcminute of the line of sight the $\pr(\kappa)$ would be almost as good as having complete spectroscopic coverage. High $\kappa$ lines of sight benefit the most from additional spectral coverage, since these LoS have the most uncertain $\kappah$. Consistent with our previous findings the lowest $\kappa$ lines of sight have the most constrained $\kappa$ PDFs.

@@

\section{Testing for Systematic errors}
\label{sec:biases}
\begin{figure*}
\includegraphics[width=\textwidth]{figs/biasplots.eps}
\caption[magcut]{Estimating $\kappa-\kappa_{\rm true}$ from subsets of
lenses. By subtracting the true (ray-traced) convergence from the
infered P($\kappa$) for each line of sight we are left with a PDF that
should be estimating zero, these can be multiplied together for multiple
lines of sight to test for a possible bias. These plots show the
expectation value of $\prod_{i=1}^N \pr_i(\kappa-\kappa_{\rm true})$ -
deviations from zero can be seen as a bias. The solid, dashed and dotted
lines are for combining 20, 5 and 2 lenses respectively. Green lines are
from multiplying P($\kappa$) \infered from photometry alone, whilst red
lines are from multiplying P($\kappa$) \infered from photometry and a
spectroscopic redshift for all halos with $i<23$ and all halos within 1
arcminute and $i<24$. The left most plot shows randomly selected lines
of sight. The central plot shows sight lines randomly selected from the
third of lines whose P($\kappa$) is most tightly constrained. The right
hand plot shows only lines of sight with external shear of 0.03 or
greater.}
\label{fig:biasplots}
\end{figure*}

Whilst it is good to reconstruct $\kappa$ precisely it is vital that
$\kappa$ is reconstructed accurately. A precise, inaccurate $\kappa$
estimator is worse than useless for cosmology. We test the accuracy of
our reconstructed $\pr(\kappa)$ in the following way: for each line of
sight we shift the reconstructed $\pr(\kappa)$ by the line of sight's
raytraced $\kappa$ to give $\pr(\kappa-\kapparay)$. Since the
$\pr(\kappa-\kapparay)$ for each line of sight is an \independant estimator of zero, the PDFs can be combined according to

\be
\label{eq:bias}
\mathcal{P}_N=\prod_{i=1}^N \pr_i(\kappa-\kappa_{\rm true})
\ee

We quantify the bias as the size of the deviation in the expectation value of $\mathcal{P}$ from zero. If the bias of $\mathcal{P}_N$ is typically smaller than the half-width of $\mathcal{P}_N$ our $\kappa$ estimators can be considered accurate. The number of useful time delay lines of sight can be approximated as the point at which the modulus of the bias and width of $\mathcal{P}_N$ are the same size.

%\todo{tom}{do}
%\subsection{Using the wrong M$_{\rm halo}$-M$_{*}$ relation}
\subsection{Selection bias from a subset of lines of sight?}

First we test for systematic errors in the reconstruction of a randomly
selected subset of samples. With the purely photometric reconstruction
there is a small bias towards underestimating $\kappa$,  combining 6
lenses the half-width and bias of $\mathcal{P}_{6}$ both equal 0.007.
With targeted spectroscopic follow up this improves slightly, for the
combination of 8 lenses the half-width and expectation value 
$\mathcal{P}_{8}$ equal 0.005. Since strong lens modelling uncertainties
are typically at the 3\% level, several hundred lenses are needed to
reach this level of precision, but at that point a 0.005 bias in
$\kappa$ estimation would become the limiting factor on time-delay
cosmography.

\subsubsection{Selecting only the lines of sight with tight $\pr(\kappax)$}
\label{sec:tightPDF}

Since lens monitoring campaigns are expensive, it is likely that only a
subset of lenses will be observed. By pre-selecting the lenses with the
most well constrained $\pr(\kappax)$, the cosmological value per lens
can be increased, however this has the potential to induce a bias. It is
probable that only photometry will be available at the time of lens
selection (although spectroscopic follow-up might be conducted at a
later date). We mimik such a selection by drawing lines of sight only
from those in the lowest third of $\sigma_{\kappa}$ given a photomtric
reconstruction of their fields. We find that selecting the most tightly
constrained lines of sight {\it decreases} the bias - our reconstruction
is most succesful for the lowest $\kappa$ lines of sight. With
photometry alone the bias is at the 0.0028 level. With a targeted
spectroscopic follow-up of the selected lines of sight there is
negligible improvement, since the best constrained lines of sight are
the most empty, but it is the high-$\kappa$ lines of sight that benefit
the most from spectroscopy. Photometry of the field is sufficient to
adequately select the best lines of sight for cosmology - and doing so
greatly increases both the precision and accuracy of the cosmological
constraints per lens.

\subsubsection{Selecting only high shear lines of sight}

\todo{phil}{Write up why lenses are likely to be gamma biased}

\notes{The uncertainties from modeling a strong lens system depends on
the comfiguration of the source images. With only two point images ofthe
source the lens slope and mass are degenerate; the mass-slope
degeneracy\citep[need citation][Saha?]{mass-slope degeneracy}. Where
lensed arcs or four point images are present the mass-slope degeneracy
can be broken and cosmology can be extracted, \citep{Suyu2012a}. The
detectability of arcs depend on the source, but the presence of four
images (so called quads) depend on the shear of the lens model - a high
shear is needed to produce a quad. The shear is almost certainly
dominated by the lens model itself, but it is possible that quads are
biased towards high external shear.}

By selecting only lines of sight with an external shear of $\gamma>0.03$ we can test the effect of shear selection. We find that in all cases $\kappa$ is underestimated at the $\sim$0.01 level. Part of this effect is mitigated if only the lines of sight with both $\gamma>0.03$ and sharp $\pr(\kappax)$ are used but biases at the $\sim$0.005 level remain. Since our model does not include shear constraints it is not surprising that selecting based on shear can induce a bias. A more sophisticated model that includes the shear recovered from the lens modelling may not be susceptable to this bias. The toy model of this section ($\gamma>0.03$) is likely to be a much stronger selection than reality, but it is worth noting that the reconstruction \proceedure will be biased if lens selection functions are extreme and unaccounted.

@@

We have shown that reconstructing the matter due to halos along any line
of sight can improve the constraints on the external convergence along
that line of sight. The total convergence along a line of sight is
strongly correlated with the reconstructed $\kappah$. However since our
model ignores voids and assumes all halos follow a spherical
truncated-NFW profile our model does not include all of the relevant
physics, hence the width of our resulting $\pr(\kappax)$ is still
typically $\sim$0.01 for any given lightcone, even with a perfect
knowledge of every halo's virial mass and redshift. To make further
progress a more advanced treatment of both voids and halos will be
necessary. Interestingly, we have found that the most empty lines of
sight can be reconstructed with the most precision. $\kappax$ for empty
lines of sight have little contributions from halos and a large
contribution from voids; since they have the tightest PDFs after the
reconstruction it seems that our model's biggest uncertainties are
driven by naively reconstructing halos rather than neglecting voids.
With a photometric reconstruction of the field there is a significant
broadening but the resulting precision is still typically 1.5 times
greater than using the global $\kappa$ distribution.

Since the uncertainty on $\pr(\kappa|\mathcal{D})$ is $\sim$0.01 even
with perfect knowledge of halo mass and redshift it seems that voids
(plus deviations from sphericity and dark matter clumping within the
main halo) dominate the total $\kappa$ uncertainty even with very
uncertain stellar masses. Inferring halo ellipticity and dark matter
clumping will likely always remain a difficulty for line of sight
reconstruction; but as Figure \ref{fig:magcut} shows observing deeper
than 24th magnitude does not signifcantly help the reconstruction.
Because our reconstruction is mostly limited by the model, spectroscopic
coverage can only provide a modest improvement to the reconstruction and
at significant observational cost.

For a small fraction of lines of sight, $\pr(\kappax)$ remains very
broad even after applying our reconstruction; these lines of sight are
typically the most overdense lines of sight in the universe. For
time-delay cosmography the most observationally expensive task is the
lightcurve monitoring, whilst making photometric observations of a
4$\times$4 arcminute patch of sky survey down to 24th magnitude is a
relatively cheap. We have shown that with a single epoch observation of
the region around a strong lens it is possible to infer $\pr(\kappax)$.
If a line of sight has a broad $\pr(\kappax)$ it can be rejected {\it
before} the investment of longterm lightcurve monitoring. We find that
selecting only the lines of sight with the most tightly constrained
$\pr(\kappax)$ does not induce a bias.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\includegraphics[width=\columnwidth]{figs/biasplots3.eps}
\caption{}
\label{fig:no-median-tricks}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The method we have outlined can also be used to estimate the external shear
along a line of sight. Shear is an observable that can be extracted from
strong lens modelling, however there is a degeneracy between internal and
external shear. Progress has been made in disentangelling external and
internal shear \citep[\eg][]{xxx} but there are still significant
uncertainties: \citet{WongEtal2011} attempted to match the shear from strong
lens models with a reconstruction of the local lens group environment, but
found a tension between the strong lens model and the reconstruction of the
environment. Since \citet{WongEtal2011} only reconstructed the local lens
group, rather than the full line of sight contribution, it is unclear whether
the external shear from lens models can be reconciled with a line of sight
reconstruction. {\it If} external shear can be measured, it provides an
additional constraint on which of the Millenium Simulation lines of sight are
similar to the reconstructed line of sight. \citet{SuyuEtal2012} found that
for RXJ1131 combining shear constriants with galaxy number count overdensity,
gave a significantly different $\pr(\kappax|\gamma,N_{45})$ copmpared to the
PDF from number count overdensity alone, $\pr(\kappax|N_{45})$. Whilst not included
in this paper, we found that given perfect knowledge of the halo mass and redshift
the raytraced external shear $\gamma$ and the reconstructed external shear $\gamma_{\mathrm{h}}$ 
are similar, with 68 percent of lines obeying
\be
\label{eq:shearineq}
|{\boldmath{\gamma}-\boldmath{\gamma_{\mathrm{h}}}}| < \mathrm{0.025}
\ee 
Future work should investigate whether $\gamma_{\mathrm{h}}$ can be used
to improve the accuracy and precision of $\kappa$ estimation, given a
reconstruction of the line of sight.

This work has used the \MS as a calibration guide. If the real universe
is not like  the \MS the the calibration will be incorrect - repeating
the study with a different simulation would quantify the sensitivity of
our results to the simulation used. Similarly the use of a different
semi-analytic model to paint galaxies onto halos could provide a useful
check, to ensure against potential systematic errors.

>>>>

\author[Collett \etal]{%
  Thomas~E.~Collett$^{1}$\thanks{\collettemail},
  Philip~J.~Marshall$^{2}$,
  Matthew~W.~Auger$^{1}$,
  Stefan~Hilbert$^{3}$,
\newauthor{%
  Sherry~H.~Suyu$^{4}$,
  Zachary~Greene$^{4}$,
  Tommaso~Treu$^{4}$\thanks{\packard},
  Vasiliy~Belokurov$^{1}$,}
\newauthor{%
  Christopher~D.~Fassnacht$^{5}$,
  L\`eon~V.~E.~Koopmans$^{6}$,
  Roger~D.~Blandford$^{3}$} 
  \medskip\\
  $^1$\ioa\\
  $^2$\oxford\\
  $^3$\kipac\\
  $^4$\ucsb\\
  $^5$\davis\\
  $^6$\kapteyn
}

@@

We now move on to consider the halo model reconstruction of line of sight mass
distributions given noisy astronomical observables. A typical imaging survey
can be expected to provide measurements of the positions and magnitudes of
galaxies in a field;  spectra for some of the objects may either come from a
synergistic survey, or from targeted follow-up. In this section we quantify
the errors induced by inferring the halo mass and redshift from these
observables.

As described in \Sref{sec:model:halos}, in this work we infer halo masses
given measurements of stellar mass. We will investigate two main sources of
error: inferring halo mass given observed stellar mass, and placing halos at
the wrong redshift due to photometric redshift error. Much work has already
focused on using photometric colours to infer stellar mass
\citep[\eg][]{AugerEtal2009} and redshifts \citep[\eg][]{BPZ}: we will
estimate the likely uncertainties on stellar mass and redshift based on this
work.

\subsection{Making Mock Observational Catalogs}

While the \MS catalogs do already contain stellar masses for each galaxy, we
find that these are not well behaved... \todo{Tom}{Explain why we re-made the
stellar masses}. We also wanted to be able to perform the functional test of
trying to recover the convergence having assumed the correct stellar
mass--halo mass relation. For these reasons, we assign a new stellar mass to
each halo in the \MS catalogs, according to the stellar mass--halo mass
relation of  \citet{BehrooziEtal2010}. From this true  stellar mass we
simulate an observed stellar mass by drawing a sample from 
$\pr(\log{M^{*}_{\rm obs}}|\log{M^{*}_{\rm true}})$ which we take to be a
Gaussian of width $\sigma_{M_*}$ centred on $\log(M*_{\mathrm {true}})$. Where
a spectroscopic redshift exists stellar masses can be estimated with typical
uncertainties of 0.15 dex \citep{AugerEtal2009}; however with photometric
redshifts stellar mass uncertainties are typically three times as large. We
use $\sigma_{M_*}=0.15$ for halos with a spectroscopic redshift and
$\sigma_{M_*}=0.45$ otherwise. For photometric redshift errors we draw a
redshift from $\pr(z_{\rm true}|z_{\rm obs})$ which we take to be a Gaussian
of width $0.1(1+z_{\rm spec})$ centred on $z_{\rm spec}$, where $z_{\rm spec}$
is the halo's true redshift in the \MS catalogue.

@@

Given an uncertain stellar mass and redshift it is possible to infer a
halo mass using the stellar mass--halo mass relation of
\citet{BehrooziEtal2010}. This \proceedure requires an inversion of the
relation given in \citet{BehrooziEtal2010} and correctly inverting the
relation's uncertainties requires care: the \proceedure we use to do this
is given in \Aref{appendix:MSMH}. By drawing sample halo masses
and redshifts, we can infer a sample $\kappah$ using the \proceedure of
\Sref{sec:model}; repeatedly drawing samples allows us to
estimate $\pr(\kappah|\data)$ for each reconstructed line of sight. We then
transform this into our target PDF $\pr(\kappa|\data)$.

\begin{figure}
\includegraphics[width=\columnwidth]{figs/Width.eps}
\caption{The widths of the \infered PDFs $\pr(\kappah|\data)$ for
$10^5$ lines of sight, given different quality of data. Blue:
spectroscopic redshift for every halo with $i<26$; Purple: 
spectroscopic redshift for every halo with $i<23$, and every halo with $i<24$
within 1 arcminute; Cyan: spectroscopic redshift for every halo with $i<22$; 
Green: all the objects in the field have only photometric redshifts.}
\label{fig:reconwidths}
\end{figure}

\begin{figure}
\includegraphics[width=\columnwidth]{figs/Widthvshilberrt.eps}
\caption{Width of the \infered PDF $\pr(\kappah|\data)$ versus the true
$\kappa$ along a line of sight. Blue assumes a spectroscopic redshift for
every halo, whilst green assumes only photometric redshifts. The width of the
\infered $\pr(\kappah|\data)$ increases with $\kappa$. 
\todo{Tom}{Check that you mean
$\kappah$ and not $\kappa$, and potentially change the graph!}}
\label{fig:widthsvsH}
\end{figure}

The best possible reconstructions of $\kappa$ will come from having a
spectroscopic redshift for every single object in each field
(\Fref{fig:reconwidths}).  In terms of telescope time, such a reconstruction
would be prohibitively expensive, but we investigate this scenario as an ideal
case. Reconstructing the lines of sight with perfect knowledge of the redshift
but an uncertain stellar mass, we find that the width of $\pr(\kappa|\data)$
grows with the expectation value of $\kappa$ (\Fref{fig:widthsvsH}). The low
$\kappa$ lines of sight are relatively empty, and so there are few
opportunities for uncertainties in the halo masses to \propogate into $\kappa$
uncertainties. The median reconstruction width in this idealised situations is
0.0123; some 22\% of lines of sight would have a reconstruction width less
than 0.01.

% PJM: Note that I removed talk of void correction here - surely we should be
% just applying the method we describe and inferring $\kappa$?

@@

Inferring the stellar mass of a galaxy from its magnitude and colours requires
an estimate of how far away the galaxy is; without a spectroscopic redshift
the \infered stellar mass is less precise. However, obtaining photometry has
much lower observational cost; over the next few years several large area
photometric surveys will reach 24th magnitude \citep{Euclid,LSST}, these will
provide sufficient data to reconstruct lines of sight {\it without} additional
observations. In principle the photometric redshift is correlated with the
\infered stellar mass, however we do not model this effect since the
convergence from the outskirts of an individual halo is only weakly \dependant
on redshift at fixed mass: the redshift error has only a small effect on the
inferred $\pr(\kappa|\data)$ when compared to the effect of uncertain stellar
masses.  With only photometric redshifts the uncertainty on
$\kappah$ is much larger than in the spectrocopic case and this
\propogates into a broader $\pr(\kappa|\data)$, although the photometric
reconstruction still typically produces a 40\% improvement compared to the
precision of the global $\pr(\kappa)$ (\Fref{fig:widthsvsH}).

@@

While a fully photometric reconstruction provides useful constraints on
$\pr(\kappa|\data)$, targeted spectroscopy can provide addtional constraints
on the masses and redshifts of halos whose $\kappah$ contribution have the
largest absolute uncertainty. Obtaining spectra of bright ($i<22$) objects is
relatively fast with large telescopes: however, we find that if spectroscopic
redhsifts were known for all $i<22$ galaxies in our field the reconstruction
improves slightly upon the purely photometric reconstruction, although the
improvement can depend strongly on the details of the particular line of
sight. Obtaining spectra for fainter objects would be correspondingly more
expensive, but if spectroscopic redshifts could be obtained for all $i<23$
galaxies and all $i<24$ galaxies within 1 arcminute of the line of sight, the
$\pr(\kappa|\data)$ would be almost as good as having complete spectroscopic
coverage (\Fref{fig:reconwidths}). High $\kappa$ lines of sight benefit the
most from additional spectral coverage, since these LoS have the most
uncertain $\kappah$. Consistent with our previous findings the lowest $\kappa$
lines of sight have the most constrained $\kappa$ PDFs.

@@

\section{Systematic Errors due to Sample Selection}
\label{sec:bias}

\begin{figure*}
\includegraphics[width=\textwidth]{figs/biasplots.eps}
\caption{Estimating $\kappa-\kappa_{\rm true}$ from subsets of
lenses. By subtracting the true (ray-traced) convergence from the
infered P($\kappa$) for each line of sight we are left with a PDF that
should be estimating zero, these can be multiplied together for multiple
lines of sight to test for a possible bias. These plots show the
expectation value of $\prod_{i=1}^N \pr_i(\kappa-\kappa_{\rm true})$ --
deviations from zero constitute biases. The solid, dashed and dotted
lines correspond to combinations of 
20, 5 and 2 sightlines respectively. Green lines show the results
from multiplying $\pr(\kappa|\data)$ \infered from photometry alone, whilst red
lines are from multiplying $\pr(\kappa)$ \infered from photometry and a
spectroscopic redshift for all halos with $i<23$ and all halos within 1
arcminute and $i<24$. The leftmost plot shows randomly selected lines
of sight. The central plot shows sight lines randomly selected from the
33\% of lines whose P($\kappa$) is most tightly constrained. The right
hand plot shows only lines of sight with external shear of 0.03 or
greater.}
\label{fig:biasplots}
\end{figure*}

While it is good to reconstruct $\kappa$ precisely, it is more important that
$\kappa$ is reconstructed accurately. 
% A precise, inaccurate $\kappa$ estimator
% is worse than useless for cosmology. 
% PJM: steady on.
We test the accuracy of our reconstructed $\pr(\kappa|\data)$ in the following
way: for each line of sight we shift the reconstructed $\pr(\kappa|\data)$ by
the line of sight's true value of $\kappa$ to give
$\pr(\kappa-\kappatrue|\data)$. Since the $\pr(\kappa-\kappatrue|\data)$ for
each line of sight is then an \independant estimator of zero, the PDFs can be
combined according to
\be
\label{eq:bias}
\mathcal{P}_N = \prod_{i=1}^N \pr_i(\kappa-\kappa_{\rm true})
\ee
In this section, we quantify the bias as the size of the deviation in the
expectation value of $\mathcal{P}_N$ from zero. If the bias of $\mathcal{P}_N$
is smaller than the half-width of $\mathcal{P}_N$, our $\kappa$ estimators can
be considered accurate. 
The point at which the modulus of the bias and the width of $\mathcal{P}_N$
are approximately equal gives an estimate of the number of objects which can
be combined before the measurement is limited by systematic error.

The source of systematic error that we investigate  is that of sample
selection: we define three different selections of lines of sight, and compute
the bias present in each one in turn, as a function of data quality. The
results described below are illustrated in \Fref{fig:biasplots}.

\subsubsection{Selecting random lines of sight}
\label{sec:bias:random}

We first consider a purley random selection of lines of sight. With the purely
photometric reconstruction, there is a small bias towards underestimating
$\kappa$: combining 6 targets the half-width and bias of $\mathcal{P}_{6}$
both equal 0.007. With targeted spectroscopic follow up this improves
slightly, for the combination of 8 sightlines the half-width and expectation
value  $\mathcal{P}_{8}$ equal 0.005.  If these lines of sight hosted time
delay lenses, the latter number would correspond to a systematic error of
0.5\% in time delay distance. Since strong lens modelling uncertainties are
typically at the 3-4\% level, a sample of 8 lenses would still be limited by
modelling uncertainty. However, a sample containing 100 or more lenses or more
would be limited by the external convergence reconstruction instead.

\subsubsection{Selecting only the lines of sight with tight $\pr(\kappax)$}
\label{sec:bias:tightPDF}

Since follow-up campaigns (such as high resolution imaging, or time
variability monitoring) are expensive, it is likely that only a subset of
detected objects will be observed further. By pre-selecting objects that have 
the most well constrained convergence, the cosmological value per lens can be
increased. However, this selection has the potential to induce a bias. It is
probable that only photometry will be available at the time of sub-sample
selection (although spectroscopic follow-up might be conducted at a later
date). We mimic such a selection by drawing lines of sight only from those in
the lowest third of $\sigma_{\kappa}$, given a purely photometric
reconstruction of their fields.

We find that selecting the most tightly constrained lines of sight {\it
decreases} the bias -- our reconstruction is most \succesful for the lowest
$\kappa$ lines of sight. With photometry alone, the bias in this sample is
0.0028. With targeted spectroscopic follow-up of the selected lines of sight
there is negligible improvement, since the best constrained lines of sight are
the most empty, and it is the high-$\kappa$ lines of sight that benefit the
most from spectroscopy. Photometry of the field is sufficient to adequately
select the best lines of sight for cosmology -- and doing so greatly increases
both the precision and accuracy of the cosmological constraints per system.

\subsubsection{Selecting only high shear lines of sight}
\label{sec:bias:tightPDF}

Time delay lenses are often selected for follow-up based on their images'
brightness (to make monitoring observations less expensive), their images'
separation (to decrease the covariance between the ground-based image
lightcurves), or the fact that they are quadruply-imaging rather than
doubly-imaging, since this yields 3 time delays instead of 1, and a higher
magnification, more informative Einstein Ring system. All three of these
properties favour lenses with high convergence and shear along their lines of
sight. Focusing on the quad selection, we might expect many of these systems
to have significant external shear (since the cross-section for 4 image
production is so sensitive to ellipticity in the lens mass distribution (or
equivalently, external shear due to the lens' environment).

By selecting only lines of sight with an external shear of $\gamma>0.03$ we
can test the impact of shear selection. We find that in this sample, $\kappa$
is systematically underestimated at the $\sim$0.01 level. Part of this effect
is mitigated if only the lines of sight with both $\gamma>0.03$ and sharp
$\pr(\kappax)$ are used but biases at the $\sim$0.005 level remain. Since our
model does not include shear constraints it is not surprising that selecting
based on shear can induce a bias. A more sophisticated model that includes the
shear recovered from the lens modelling may be less susceptable to this bias.
The toy model of this section ($\gamma>0.03$) is likely to be a much stronger
selection than reality, but it is worth noting that the reconstruction
\proceedure will be biased if lens selection functions are extreme and
unaccounted for.

@@

We have shown that reconstructing the matter due to halos along any line
of sight can improve the constraints on the external convergence along
that line of sight. The total convergence along a line of sight is
strongly correlated with the reconstructed $\kappah$. However since our
model ignores voids and assumes all halos follow a spherical
truncated-NFW profile our halo model does not include all of the relevant
physics; hence, the width of our resulting $\pr(\kappa|\data)$ is still
typically $\sim$0.01 for any given lightcone, even with a perfect
knowledge of every halo's virial mass and redshift. To make further
progress a more advanced treatment of both voids and halos will be
necessary. Interestingly, we have found that the most empty lines of
sight can be reconstructed with the most precision. $\kappa$ estimates
for empty
lines of sight receive few contributions from halos, and a large
contribution from voids; since they have the tightest PDFs after the
reconstruction it seems that our model's biggest uncertainties are
driven by naively reconstructing halos rather than neglecting voids.
With a photometric reconstruction of the field there is a significant
broadening, but the resulting precision is still typically 50\% 
higher than using the global $\kappa$ distribution.

Since the uncertainty on $\pr(\kappa|\data)$ is $\sim$0.01 even
with perfect knowledge of halo mass and redshift, it seems that voids
(plus deviations from sphericity and dark matter clumping within the
main halo) dominate the total $\kappa$ uncertainty even with very
uncertain stellar masses. Inferring halo ellipticity and dark matter
clumping will likely always remain a difficulty for line of sight
reconstruction; but as \Fref{fig:magcut} shows, observing deeper
than 24th magnitude does not signifcantly help the reconstruction.
Because our reconstruction is mostly limited by the model, spectroscopic
coverage can only provide a modest improvement to the reconstruction and
at significant observational cost.

For a small fraction of lines of sight, $\pr(\kappa|\data)$ remains very
broad even after applying our reconstruction; these lines of sight are
typically the most overdense lines of sight in the universe. For
time-delay cosmography the most observationally expensive tasks are the
lightcurve monitoring and high resolution follow-up imaging and spectroscopy,
while making photometric observations of a
4$\times$4 arcminute patch of sky survey down to 24th magnitude is
relatively cheap. We have shown that with a single epoch observation of
the region around a strong lens it is possible to infer $\pr(\kappa|\data)$.
If a line of sight has a broad $\pr(\kappa|\data)$ it can be rejected {\it
before} the investment of follow-up. We find that
selecting only the lines of sight with the most tightly constrained
$\pr(\kappa)$ does not induce a bias in $\kappa$, and hence distance.

The method we have outlined can also be used to estimate the external shear
along a line of sight. Shear is an observable that can be extracted from
strong lens modelling, however there is a degeneracy between internal and
external shear. When the Einstein Ring data are very good it is possible to
disentangle external and internal shear \citep[\eg][]{xxx}, but there are
still significant uncertainties: \citet{WongEtal2011} attempted to match the
shear from strong lens models with a reconstruction of the local lens group
environment, but found a tension between the strong lens model and the
reconstruction of the environment. Since \citet{WongEtal2011} only
reconstructed the local lens group, rather than the full line of sight
contribution, it is unclear whether the external shear from lens models can be
reconciled with a line of sight reconstruction. {\it If} external shear can be
measured, it provides an additional constraint on which of the \MS lines of
sight are similar to the reconstructed line of sight. \citet{SuyuEtal2012}
found that for RXJ1131 combining shear constriants with galaxy number count
overdensity, gave a significantly different $\pr(\kappa|\gamma,N_{45})$
compared to the PDF from number count overdensity alone,
$\pr(\kappax|N_{45})$. While not included in this paper, we found that given
perfect knowledge of the halo mass and redshift the raytraced external shear
$\gamma$ and the reconstructed external shear $\gamma_{\mathrm{h}}$  are
similar, with 68 percent of lines obeying
\be
\label{eq:shearineq}
|{\boldmath{\gamma}-\boldmath{\gamma_{\mathrm{h}}}}| < \mathrm{0.025}
\ee 
Future work should investigate whether $\gamma_{\mathrm{h}}$ can be used
to improve the accuracy and precision of $\kappa$ estimation, given a
reconstruction of the line of sight.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\includegraphics[width=\columnwidth]{figs/biasplots3.eps}
\caption{}
\label{fig:no-median-tricks}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The reduction of $\pr(\kappah|\data)$ to just its median value, before using
it to look up the appropriate $\pr(\kappa)$ from the simulated sightline
ensemble, is reminiscent of the procedure followed by \citet{SuyuEtal2010},
and Greene et al (in prep.) The reconstruction method can be seen as the
limiting case of the weightings explored by Greene et al: rather than using
weighted number counts in an aperture, we treat each galaxy individually, and
predict  $\kappa_{\rm h,obs}^{\rm med}$ for use in their place. One
significant disadvantage of these types of method is that they impose a
bottleneck in the flow of information. Once you have compressed to one number,
it is difficult to incorporate more information later; you have to recompute
all the calibrations. Using the full PDF for $\kappah$ is the more correct
thing to do: at present, however, it suffers from significant bias in the
photometric data case. \Fref{fig:no-median-tricks} shows how $\kappa$ is
overestimated, as a result of the high scatter in halo mass for objects with 
high stellar mass. Future work should increasing the precision of the halo
mass estimates.
